\documentclass[a4paper,11pt]{article}
% --- Packages ---
\usepackage{amsmath, amssymb, textcomp, enumitem, mathptmx, hyperref, xcolor, pifont}

% --- Hyperref settings ---
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    citecolor=blue,
    urlcolor=blue
}

% --- Page geometry ---
\usepackage[a4paper,
    top=0.75in,
    bottom=0.75in,
    left=0.65in,
    right=0.65in,
    headheight=14pt,
    headsep=0.2in,
    footskip=0.4in
]{geometry}

% --- No indentation ---
\setlength{\parindent}{0pt}

% --- Bullet itemization ---
% \renewcommand{\labelitemi}{\textopenbullet}
% \renewcommand{\labelitemii}{\textbullet}
\setlist[itemize]{noitemsep, nolistsep, leftmargin=2.5em}

\usepackage{marvosym}
\usepackage{twemojis}
\usepackage[explicit]{titlesec}
\usepackage{xhfill}

\def\bull{\vrule height 0.8ex width .7ex depth -.1ex }

% --- Font and ATS setup ---
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[scaled=0.9]{helvet}
\renewcommand\familydefault{\sfdefault}


% --- Custom commands ---
\renewcommand{\hrulefill}{%
\leavevmode\leaders\hrule height 0.3pt\hfill\kern0pt }

\newcommand{\area} [2] {
\vspace*{-9pt}
\begin{verse}
\textbf{#1}   #2
\end{verse}
}

\newcommand{\lineunder} {
\vspace*{-8pt} \\
\hspace*{-18pt} \hrulefill \\
}

\newcommand{\header} [1] {
\vspace{1mm}
    {\textsc{\textbf{\Large{\xrfill[0.5ex]{0.5pt}~#1~\xrfill[0.5ex]{0.5pt}}}}}
}

\newcommand{\employer} [3] {
    { \textsc{#1} (#2)\\ \underline{\textsc{\textit{#3}}}\\  }
}

\newcommand{\contact} [3] {
\vspace*{-10pt}
\begin{center}
        {\Huge \scshape {#1}}\\
        #2 \\ #3
\end{center}
\vspace*{-8pt}
}

\newenvironment{achievements}{
\begin{list}
        {$\bullet$}{\topsep 0pt \itemsep -2pt}}{\vspace*{4pt}
\end{list}
}

\newcommand{\schoolwithcourses} [4] {
\textsc{#1} #2 $\bullet$ #3\\
    #4 \\
\vspace*{5pt}
}

\newcommand{\school} [4] {
\textsc{#1} #2 $\bullet$ #3\\
    #4 \\
}

\begin{document}
\vspace*{-0.45in}
\fontsize{10.0}{11.5}\selectfont

\begin{raggedright}
    {\Huge\textbf\textsc{{{Don Lim}}}} \hfill
\href{mailto:lim.don93@gmail.com}{lim.don93@gmail.com} | (806) 317-7937 | \href{https://donlim93.github.io/}{https://donlim93.github.io} | \href{https://www.linkedin.com/in/lim-don/}{https://www.linkedin.com/in/lim-don} \\ \vspace{2mm}
\end{raggedright}

\vspace{1mm}

Financial Systems Analyst with expertise in financial planning platforms, business analytics, and process automation. Specializes in IBM Planning Analytics (TM1), Salesforce CRM, SQL, and Python to drive forecasting accuracy, scalable reporting, and cross-functional efficiency. Proven track record of modernizing legacy Excel and SSRS workflows into governed, automated systems. Background spanning IT and finance, supported by graduate-level training in econometrics and optimization.

\vspace{1mm}
\header{Work Experience} 

\vspace{1mm}
\textbf{Green Brick Partners, Inc.} \hfill Plano, TX \\
\textit{Senior Financial Systems Analyst, Finance and Operations} \hfill \textit{Nov 2024 --- Oct 2025} 
\begin{itemize}
\item Built and automated TM1 TurboIntegrator (TI) processes and business rules, integrating first- and third-party tools (PAW, Excel, FileZilla, Cognos Command Center, Arc, Pulse, TM1py, Git), enabling daily forecasting cycles instead of biannual forecasts.
\item Programmatically maintained a TM1 OLAP database consisting of 42 cubes, 88 dimensions, attributes, subsets, control objects, security, applications, and views across development, UAT, and production environments using MDX, SQL (joins, CTEs, cross applies), TIs, and TM1 rules.
\item Partnered with the VP of Asset Management to develop a land cash flow module incorporating uniform, annual, and quarterly cost-to-complete spreads, retainage, and payout schedules, reducing forecast preparation time by 50\%.
\item Converted mortgage forecasting models from Excel to TM1, improving calculation speed and increasing granularity from business unit to project level; collaborated with the Controller to integrate personnel, headcount, and loan assumptions including closings, funded loans, loan types, application rates, contract rates, fees, and expenses.
\item Designed dynamic, auto-refreshing dashboards and executive-level visualizations for financial managers, the CFO, and board members, replacing SSRS reports and reducing board package preparation time by 75\%.
\item Implemented and managed a Git-based version control repository for TM1 processes and rules, enabling backup restoration and supporting SOX and ITGC audit compliance through automated Git Bash audit scripts.
\item Troubleshot and optimized Excel-based forecasting templates using pivot tables, VLOOKUP/XLOOKUP, dynamic array formulas, nested functions, and Power Query to resolve calculation and refresh issues.
\item Designed and automated ETL workflows for P\&L, SGM, GL, KPIs, statistical, and budget data, improving accuracy and efficiency in cash flow, KPIs, and milestone forecasting.
\item Oversaw integrations between Workiva, land proforma systems, and TM1, including API configuration and database table mappings.
\item Developed a proof of concept exporting TM1 views to Power BI dashboards using Python (TM1py), REST APIs, and DAX to enable dynamic reporting.
\item Maintained comprehensive documentation and change logs for all TM1 processes, rules, and applications to support governance, audits, and team onboarding.
\end{itemize}

\vspace{1mm}
\textit{Sales and Marketing Analyst, IT} \hfill  \textit{May 2023 --- Nov 2024} 
\begin{itemize}
\item Administered Salesforce CRM, Marketing Cloud, and Sales Cloud environments; implemented and tested enhancements across dev, UAT, and production.
\item Created Salesforce dashboards, reports, workflows, and automated campaigns, improving visibility for 200+ end users.
\item Cleaned, validated, and deduplicated Salesforce data using regex, SOQL, and Data Loader, increasing data accuracy.
\item Served as system administrator for ERP and business tools (Constellation NEWSTAR, BuildPro, Calendly, DocuSign).
\item Trained and provided life-cycle support for end users on Salesforce and business applications.
\item Integrated new prospect data from third-party lead generation tools (Zillow, Realtor.com) into Salesforce using APIs and middleware (Zapier).
\item Implemented prospect integration between Calendly and Salesforce using Flow and URL slugs, automating meeting scheduling and follow-up tasks.
\end{itemize}

\vspace{1mm}
\textbf{DocuNav Solutions} \hfill Frisco, TX \\
\textit{Support Engineer} \hfill  \textit{Jun 2022 --- May 2023} 
\begin{itemize}
\item Built, tested, and deployed Laserfiche document management solutions (Quick Fields, Workflow, Forms) that automated document OCR, extraction, routing, and classification for corporate, legal, and school district clients.
\item Maintained non-clustered SQL servers (stored procedures, task scheduling, SSL/TLS configuration).
\item Managed document- and user-level security via Active Directory and admin consoles.
\item Provided IT support for workstations, laptops, and peripheral devices.
\end{itemize}    

\vspace{1mm}
\header{Education} 

\vspace{1mm}
\textbf{Texas Tech University, Department of Agricultural and Applied Economics}\hfill Lubbock, TX\\
\textit{Master of Science in Agricultural and Applied Economics, GPA: 3.6/4.0} \hfill  \textit{May 2022}
\begin{itemize}
\item Relevant coursework: Causal Inference, Optimization, Econometrics I \& II, Production Economics, Machine Learning, Macro/Microeconomics.
\item Tools and methods: R (tidyverse), Stata, MATLAB, Python, SAS, \LaTeX; econometric methods including ordinary least squares, instrumental variables, difference-in-difference, regression discontinuity design, matching, k-Nearest Neighbors, clustering.
\end{itemize}

\vspace{1mm}
\textbf{CEVRO Institute, School of Political Studies}\hfill Prague, Czech Republic\\
\textit{M.A. in Political Science, Philosophy, and Economics (PPE)} \hfill  \textit{Jun 2019}\\
\textbf{Scholarship:} Full Tuition for Academic Excellence.

\vspace{1mm}
\textbf{University of Michigan, College of Literature, Science, and the Arts} \hfill Ann Arbor, MI\\
\textit{Bachelor of Arts in General Studies (Political Science, Philosophy, and History)} \hfill  \textit{Apr 2016}

\vspace{1mm}
\header{Research Experience} 

\vspace{1mm}
\textbf{Free Market Institute, Texas Tech University} \hfill Lubbock, TX\\
\textit{Economic Research Assistant} \hfill  \textit{Aug 2020 --- May 2022} 
\begin{itemize}
\item Conducted economic literature reviews and statistical analyses for academic publications and book chapters.
\end{itemize}

\vspace{1mm}
\textbf{American Institute for Economic Research} \hfill Great Barrington, MA\\
\textit{E.C. Harwood Visiting Research Fellow} \hfill  \textit{Dec 2018 --- May 2019}

\vspace{1mm}
\textbf{Free Market Institute, Texas Tech University} \hfill Lubbock, TX\\
\textit{Visiting Research Scholar} \hfill  \textit{Sep 2018 --- Dec 2018}
\begin{itemize}
\item Researched and presented master’s thesis on the economic organization of open-source communities.
\end{itemize}

\vspace{1mm}
\textbf{Commonwealth Foundation} \hfill King of Prussia, PA\\
\textit{Research Fellow} \hfill  \textit{Jan 2017 --- May 2017}
\begin{itemize}
\item Designed visual reports of Pennsylvania’s tax and expenditure data to support policy publications.
\item Authored and edited weekly newsletters on state budget and education reform.
\item Responded to legislative and public inquiries on fiscal and labor policy.
\end{itemize}

\vspace{1mm}
\textbf{Koch Internship Program} \hfill Online\\
\textit{Intern} \hfill  \textit{Jan 2017 --- May 2017}
\begin{itemize}
\item In concurrence with the Commonwealth Foundation fellowship, developmental program included weekly seminars on economics, public policy, and career development.
\end{itemize}

\vspace{1mm}
\textbf{Cato Institute} \hfill Washington, DC\\
\textit{Research Intern, Center for Educational Freedom} \hfill \textit{Sep 2016 --- Dec 2016}
\begin{itemize}
\item Analyzed international education datasets (TIMSS, PISA) to evaluate policy impacts.
\item Summarized congressional proceedings on education, trade, and monetary policy.
\item Produced 200+ social media data visualizations for @HumanProgress.
\end{itemize}

\vspace{1mm}
\header{Certifications and Exams} 
\vspace{1mm}

\begin{itemize}
\item PostgreSQL for Everybody Specialization \hfill Coursera, \textit{May 2024}\\
\textit{Skills:} PostgreSQL, DBeaver, NoSQL, Elasticsearch
\item Web Applications for Everybody Specialization \hfill Coursera, \textit{Oct 2023}\\
\textit{Skills:} MySQL, PHP, JavaScript, jQuery, JSON, XAMPP
\item Applied Data Science with Python Specialization \hfill Coursera, \textit{Aug 2023}\\
\textit{Skills:} Python (regex, Pandas, scikit-learn), ML algorithms (random forest, SVC, logistic regression, k-Nearest Neighbors, Na\"{i}ve Bayes, clustering)
\item Django for Everybody Specialization \hfill Coursera, \textit{Jul 2023}\\
\textit{Skills:} Django, SQLite, HTML5, CSS, JavaScript, Bash
\item Passed: Probability Exam \hfill Society of Actuaries, \textit{May 2023}
\item Laserfiche Gold \& Platinum Certifications \hfill Laserfiche, \textit{Oct 2022}
\end{itemize}

\end{document}
