---
title: "PCCA Survey 2022"
author: "Don Lim"
date: "4/6/2022"
output:
  pdf_document:
    toc: yes
    keep_tex: true
  html_document:
    df_print: paged
    toc: yes
    theme: cerulean
    highlight: haddock
    code_folding: show
    number_sections: no
    toc_float:
      collapsed: yes
      smooth_scroll: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    tidy = TRUE,
    message = FALSE,
    tidy.opts = list(width.cutoff = 60),
    warning = FALSE
)
```

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)
library(estimatr)
library(stargazer)
library(pastecs)
library(reshape2)

library(ggpubr)
library(rstatix)
library(vtable)
```

# Data cleanup

Imports data and removes redundant first row

```{r}
survey <-
    read_excel(
        "SoGoSurvey_PCCA Grower-Owner Survey 3-25-22.xls",
        col_types = c(
            "skip",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text",
            "text"
        )
    )

survey <- survey[-c(1), ]
survey <- survey[-c(355:362), ]
```
Rename columns

```{r}
names(survey) <-
    c(
        'number',
        'status',
        'gathering information',
        'social media',
        'traditional media',
        'interpersonal communication',
        'industry organization',
        'my cooperative',
        'other',
        'new media',
        'facebook',
        'twitter',
        'instagram',
        'youtube',
        'snapchat',
        'tiktok',
        'other2',
        'podcast or radio',
        'podcast list',
        'satisfaction',
        'market information',
        'crop information',
        'industry news',
        'farm operation',
        'other3',
        'website',
        'email',
        'member access',
        'mypcca',
        'text',
        'social media2',
        'print',
        'local coop',
        'other4',
        'area pcca does best',
        'mypcca app',
        'social media3',
        'email2',
        'text2',
        'printed',
        'other5',
        'communication improvements',
        'topic',
        'accessible',
        'responsive',
        'courteous',
        'marketing pool',
        'the seam',
        'pcca direct',
        'forward contracting',
        'most used marketing method',
        'information',
        'change',
        'influence',
        'best performance',
        'minimize risk',
        'control',
        'reliability',
        'price information',
        'why member',
        'opportunities',
        'how',
        'open and voluntary',
        'member control',
        'economic participation',
        'autonomy',
        'education',
        'cooperation',
        'community',
        'coop vs noncoop',
        'knowledge of coop',
        'coop meetings',
        'up to date',
        'age',
        'description',
        'active',
        'farming duration',
        'region',
        'operation',
        'education level',
        'membership duration',
        'thoughts'
    )
```

Dummy variables

Replaces 1 if true, 0 if false

```{r}
survey <-
    mutate_at(
        survey,
        c(
            'social media',
            'traditional media',
            'interpersonal communication',
            'industry organization',
            'my cooperative',
            'facebook',
            'twitter',
            'instagram',
            'youtube',
            'snapchat',
            'tiktok',
            'market information',
            'crop information',
            'industry news',
            'farm operation',
            'website',
            'email',
            'member access',
            'mypcca',
            'text',
            'social media2',
            'print',
            'local coop',
            'mypcca app',
            'social media3',
            'email2',
            'text2',
            'printed'
        ),
        ~ replace(.,!is.na(.), 1)
    )

survey <-
    mutate_at(
        survey,
        c(
            'social media',
            'traditional media',
            'interpersonal communication',
            'industry organization',
            'my cooperative',
            'facebook',
            'twitter',
            'instagram',
            'youtube',
            'snapchat',
            'tiktok',
            'market information',
            'crop information',
            'industry news',
            'farm operation',
            'website',
            'email',
            'member access',
            'mypcca',
            'text',
            'social media2',
            'print',
            'local coop',
            'mypcca app',
            'social media3',
            'email2',
            'text2',
            'printed'
        ),
        ~ replace(., is.na(.), 0)
    )
```

```{r}
survey <- survey %>%
    mutate(across(
        c(`new media`, `podcast or radio`, `opportunities`, `active`),
        ~ factor(ifelse(.x == "Yes", 1, 0))
    ))
```

Converts Likert answers to numerical values: Strongly disagree is 1, disagree is 2, somewhat disagree is 3, somewhat agree is 4, agree is 5, and strongly agree is 6. Also changes knowledge answers to numerical values: no knowledge is 1, some knowledge is 2, and extensive knowledge is 3.

```{r}
survey <-
  mutate_at(
    survey,
    c(
      'accessible',
      'responsive',
      'courteous',
      'minimize risk',
      'control',
      'reliability',
      'price information'
    ),
    funs(
      recode(
        .,
        "Strongly Disagree" = 1,
        "Disagree" = 2,
        "Somewhat Disagree" = 3,
        "Somewhat Agree" = 4,
        "Agree" = 5,
        "Strongly Agree" = 6
      )
    )
  )

survey <-
  mutate_at(
    survey,
    c(
      'marketing pool',
      'the seam',
      'pcca direct',
      'forward contracting'
    ),
    funs(
      recode(
        .,
        "No Knowledge" = 1,
        "Some Knowledge" = 2,
        "Extensive Knowledge" = 3
      )
    )
  )
```

The following code converts character columns into numeric

```{r}
cols.num <- c(
  'number',
  'social media',
  'traditional media',
  'interpersonal communication',
  'industry organization',
  'my cooperative',
  'new media',
  'facebook',
  'twitter',
  'instagram',
  'youtube',
  'snapchat',
  'tiktok',
  'podcast or radio',
  'satisfaction',
  'market information',
  'crop information',
  'industry news',
  'farm operation',
  'website',
  'email',
  'member access',
  'mypcca',
  'text',
  'social media2',
  'print',
  'local coop',
  'mypcca app',
  'social media3',
  'email2',
  'text2',
  'printed',
  'accessible',
  'responsive',
  'courteous',
  'marketing pool',
  'the seam',
  'pcca direct',
  'forward contracting',
  'minimize risk',
  'control',
  'reliability',
  'price information',
  'opportunities',
  'open and voluntary',
  'member control',
  'economic participation',
  'autonomy',
  'education',
  'cooperation',
  'community',
  'coop vs noncoop',
  'knowledge of coop'
)

survey[cols.num] <- sapply(survey[cols.num],as.numeric)
```

## Data summary

```{r}
stat.desc(survey[cols.num])

sumtable(survey[cols.num])
```


# Data visualization

Blank theme formatting 

```{r}
blank_theme <- theme_minimal() +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(size = 14, face = "bold")
    )
```

## Age

```{r age-all}
age <-
    survey %>% group_by(age) %>% summarise(count = n_distinct(number))

bp <-
    ggplot(age, aes(x = "", y = count, fill = age)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + scale_fill_grey() + ggtitle("Age of PCCA Members") +  blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / 354, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Age group"))
```

## Area that members believe PCCA performs best at

```{r}
my.labels <- c("In-Person", "Mass Comm", "Software/Apps", "NA")
```

```{r best-all}
pcca_best <- survey %>% group_by(`area pcca does best`) %>% summarise(count = n_distinct(number))

best <- ggplot(pcca_best, aes(x = `area pcca does best`, y = count)) + geom_bar(stat = "identity") + coord_flip() + ggtitle("Area That Members Believe PCCA Performs Best") 
best + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) + scale_x_discrete(labels = my.labels)
```

## Education

```{r}
survey %>% group_by(`education level`) %>% summarise(count = n_distinct(number))
```

## Region

```{r region-all}
region <-
    survey %>% group_by(`region`) %>% summarise(count = n_distinct(number))

bp <-
    ggplot(region, aes(x = "", y = count, fill = region)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + scale_fill_grey() + ggtitle("PCCA Members Region Location") +  blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / 354, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Region"))
```

## Preferred Information

```{r preferred-all}
information <-
    survey %>% group_by(`gathering information`) %>% summarise(count = n_distinct(number))

information <- information %>%
    filter(information$count > 4)

total <- as.numeric(nrow(survey))

bp <-
    ggplot(information, aes(x = "", y = count, fill = `gathering information`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + scale_fill_grey() + ggtitle("PCCA Members Preferred Information Sources") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Types of sources"))
```

## Dry or Irrigated Operation

```{r operation-all}
operation <-
    survey %>% group_by(`operation`) %>% summarise(count = n_distinct(number))


bp <-
    ggplot(operation, aes(x = "", y = count, fill = `operation`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + scale_fill_grey() + ggtitle("Dry or Irrigated Operation") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / 354, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Types of operation"))
```

```{r meetings-all}
meetings <- survey %>% group_by(`coop meetings`) %>% summarise(count = n_distinct(number))

meetings_chart <- ggplot(meetings, aes(x = `coop meetings`, y = count)) + geom_bar(stat = "identity") + coord_flip() + ggtitle("Frequency of Attendance at PCCA Meetings") 
meetings_chart + theme(axis.title.y = element_blank(), axis.title.x = element_blank())
```


## Most used marketing method

```{r most-used-all}
marketing <-
    survey %>% group_by(`most used marketing method`) %>% summarise(count = n_distinct(number))

lbls <- paste(names(marketing), "\n", marketing, sep="")

total <- as.numeric(nrow(survey))

bp <-
    ggplot(marketing, aes(x = "", y = count, fill = `most used marketing method`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + ggtitle("Most Used Marketing Method") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Marketing methods")) 
```

```{r tenant-all}
tenant <-
    survey %>% group_by(`description`) %>% summarise(count = n_distinct(number))

total <- as.numeric(nrow(survey))

bp <-
    ggplot(tenant, aes(x = "", y = count, fill = `description`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + scale_fill_grey() + ggtitle("Tenant or Landlord") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Management"))
```


# Data analysis

## OLS

Prediction of membership satisfaction

```{r}
satisfaction = lm(satisfaction ~ accessible + responsive + courteous, data =
                           survey)
stargazer(satisfaction)
```

```{r}
satisfaction2 = lm(satisfaction ~ accessible + responsive + courteous + `marketing pool` + `the seam` + `pcca direct` + `forward contracting`, data =
                           survey)
stargazer(satisfaction2)
```

## Probit

```{r}
survey$satprob <- ifelse(survey$satisfaction >= 4, "1", "0")

survey2 <- survey %>%
  filter(!is.na(survey$satprob))

survey2$satprob <- as.numeric(survey2$satprob)
```

```{r}
myprobit <-
  glm(
    satprob ~ accessible + responsive + courteous + `marketing pool` + `the seam` + `pcca direct` + `forward contracting`,
    family = binomial,
    data = survey2
  )

stargazer(myprobit)
```


# Separating younger and older generation

In this section, I will categorize the younger generation as the age bracket 18-45 and the older generation as the age brackets for 56-65 and 65+

```{r}
younger <- survey %>% 
  filter(survey$age == "18-45")

older <- survey %>% 
  filter(survey$age == "56-65" | survey$age == "65+")

younger$gen <- "younger"
older$gen <- "older"
```

```{r}
satisfaction.y = lm(satisfaction ~ accessible + responsive + courteous, data =
                           older)
stargazer(satisfaction.y)
```

```{r best-young}
pcca_best <- younger %>% group_by(`area pcca does best`) %>% summarise(count = n_distinct(number))

best <- ggplot(pcca_best, aes(x = `area pcca does best`, y = count)) + geom_bar(stat = "identity") + coord_flip() + ggtitle("Area That Younger Members Believe PCCA Performs Best") 
best + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) + scale_x_discrete(labels = my.labels)
```

```{r best-old}
pcca_best <- older %>% group_by(`area pcca does best`) %>% summarise(count = n_distinct(number))

best <- ggplot(pcca_best, aes(x = `area pcca does best`, y = count)) + geom_bar(stat = "identity") + coord_flip() + ggtitle("Area That Older Members Believe PCCA Performs Best") 
best + theme(axis.title.y = element_blank(), axis.title.x = element_blank()) + scale_x_discrete(labels = my.labels)
```

```{r region-young}
region <-
    younger %>% group_by(`region`) %>% summarise(count = n_distinct(number))

bp <-
    ggplot(region, aes(x = "", y = count, fill = region)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

total <- as.numeric(nrow(younger))

pie + scale_fill_grey() + ggtitle("Younger PCCA Members Region Location") +  blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Region"))
```

```{r region-old}
region <-
    older %>% group_by(`region`) %>% summarise(count = n_distinct(number))

bp <-
    ggplot(region, aes(x = "", y = count, fill = region)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

total <- as.numeric(nrow(older))

pie + scale_fill_grey() + ggtitle("Older PCCA Members Region Location") +  blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Region"))
```

```{r information-young}
information <-
    younger %>% group_by(`gathering information`) %>% summarise(count = n_distinct(number))

information <- information %>%
    filter(information$count > 4)

total <- as.numeric(nrow(younger))

bp <-
    ggplot(information, aes(x = "", y = count, fill = `gathering information`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + scale_fill_grey() + ggtitle("Younger PCCA Members Preferred Information Sources") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Types of sources"))
```

```{r information-old}
information <-
    older %>% group_by(`gathering information`) %>% summarise(count = n_distinct(number))

information <- information %>%
    filter(information$count > 4)

total <- as.numeric(nrow(older))

bp <-
    ggplot(information, aes(x = "", y = count, fill = `gathering information`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + scale_fill_grey() + ggtitle("Older PCCA Members Preferred Information Sources") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Types of sources"))
```

```{r marketing-young}
marketing <-
    younger %>% group_by(`most used marketing method`) %>% summarise(count = n_distinct(number))

lbls <- paste(names(marketing), "\n", marketing, sep="")

total <- as.numeric(nrow(younger))

bp <-
    ggplot(marketing, aes(x = "", y = count, fill = `most used marketing method`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + ggtitle("Most Used Marketing Method") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Marketing methods")) 
```

```{r marketing-old}
marketing <-
    older %>% group_by(`most used marketing method`) %>% summarise(count = n_distinct(number))

lbls <- paste(names(marketing), "\n", marketing, sep="")

total <- as.numeric(nrow(older))

bp <-
    ggplot(marketing, aes(x = "", y = count, fill = `most used marketing method`)) + geom_bar(width = 1, stat = "identity")
pie <- bp + coord_polar("y", start = 0)

pie + ggtitle("Most Used Marketing Method") + blank_theme +
    theme(axis.text.x = element_blank()) +
    geom_text(aes(x = 1.7,
                  label = percent(count / total, accuracy = 0.1)),
              position = position_stack(vjust = .5)) + guides(fill = guide_legend(title =
                                                                                      "Marketing methods")) 
```

# Percentages

## Knowledge

```{r}
knowledge.cols <- c("number", "marketing pool", "the seam", "pcca direct", "forward contracting", "gen")

knowledge.younger <- younger %>% 
  select(knowledge.cols)

knowledge.older <- older %>% 
  select(knowledge.cols)
```

```{r}
knowledge.yo <- bind_rows(knowledge.younger, knowledge.older)
```

```{r}
stat.test <- knowledge.yo %>% 
  t_test(`forward contracting` ~ gen) %>% 
  add_significance()

stat.test
```









